---
title: "PA1"
output:
  html_document:
    css: custom.css
    toc: true
    toc_depth: 6
---
<br>

## Loading and preprocessing the data
Load libraries, read in data, and ensure data are in the right format
```{r}
library(ggplot2)
activityDat <- read.csv("activity.csv")
activityDat$date <- as.Date(activityDat$date, format = "%Y-%m-%d")
```
<br><br>

## What is mean total number of steps taken per day?

#### Calculate and store the total number of steps taken per day.  
```{r}
totSteps <- setNames(aggregate(activityDat$steps, 
          list(activityDat$date), sum), c("Date", "Step_Sum"))
```
<br>

#### Histogram of the total number of steps
```{r}
ggplot(totSteps, aes(x=Step_Sum)) + 
  geom_histogram(binwidth=1000, fill="white", color="black") +
  ylim(0, 20)
```
<br>

#### Mean of the total number of steps 
```{r}
mean(totSteps$Step_Sum, na.rm=TRUE)
```

#### Median of the total number of steps 
```{r}
median(totSteps$Step_Sum, na.rm=TRUE)
```

<br><br>

## What is the average daily activity pattern?
Calculate the average number of steps taken per time interval.
```{r}
avgSteps <- setNames(aggregate(activityDat$steps, list(activityDat$interval),
                               mean, na.rm=TRUE), c("Interval", "Step_Avg"))
```
<br>

#### Timeseries of average steps per time interval
```{r}
ggplot(avgSteps, aes(x=Interval, y=Step_Avg)) + geom_line()
```
<br>

#### The max # of steps across the averaged data
```{r}
avgSteps[avgSteps$Step_Avg == max(avgSteps$Step_Avg),]
```

<br><br>

## Imputing missing values

#### # of missing observations
```{r}
sum(is.na(activityDat$steps))
```
<br>

#### Devise an missing value imputation strategy
I will use the average of that 5-min interval wherever a missing value  
appears to create a new column
```{r}
activityDat$missingSteps <- avgSteps$Step_Avg
activityDat$missingSteps <- ifelse(is.na(activityDat$steps), 
        activityDat$missingSteps, 
        activityDat$steps)
```
<br>


#### Impute missing # of steps taken per day
```{r}
totStepsMissing <- setNames(aggregate(activityDat$missingSteps, 
                               list(activityDat$date), sum), c("Date", "Step_Sum_Missing"))
```
<br>

#### Histogram of sum of steps using imputed missing observations
```{r}
ggplot(totStepsMissing, aes(x=Step_Sum_Missing)) + 
  geom_histogram(binwidth=1000, fill="white", color="black") +
  ylim(0, 20)
```
<br>

#### Mean of Step_Sum_Missing
```{r}
mean(totStepsMissing$Step_Sum_Missing)
```

#### Median of Step_Sum_Missing
```{r}
median(totStepsMissing$Step_Sum_Missing)
```
The mean is the same as before, while the median is now the same as the mean.  
Furthermore, the count for the center bar is much higher now.

<br><br>

## Are there differences in activity patterns between weekdays and weekends?

#### Creating a weekday/weekend factor variable
```{r}
activityDat$day <- as.factor(ifelse(weekdays(activityDat$date) == "Sunday" | 
                            weekdays(activityDat$date) == "Saturday", 
                                   "Weekend", "Weekday"))
```
<br>

#### Calculate the average number of steps taken per time interval
```{r}
moreAvgSteps <- setNames(aggregate(activityDat$missingSteps, 
                  by=list(activityDat$interval,activityDat$day),mean),
                  c("Interval", "Day", "Step_Avg_Missing"))
```
<br>

#### Timeseries of average steps per time interval
```{r}
ggplot(moreAvgSteps, aes(x=Interval, y=Step_Avg_Missing)) + geom_line() + facet_grid(Day ~ .)
```